---
title: "Project Part 2: Data Cleaning and Rough Draft"
author: "Catalina Dorado"
format: html
execute: 
  message: false
  warning: false
---

## Part 1 - Data Load and Cleaning

**NOTE: My previous proposal on working with abalone data counts has changed due to there 
being no data available for abalones. I will now be working with Bat Star counts from two 
different sites (beaches) in Monterey and comparing their counts from 2006 to 2019.


**1. Load the appropriate R packages for analysis**

```{r load-packages}
# Loading packages
library(tidyverse)
library(stringr)
library(lubridate)
library(readr)
```


**2. Load the appropriate R packages for analysis**

```{r load-data}
# Load in the data
rcca_data <- read_csv("RCCA_invertebrate_swath_data.csv", show_col_types = FALSE)

# View data
view(rcca_data)
```


**3. Explore dataset**

```{r}
dim(rcca_data)
summary(rcca_data)
```


**4. Clean dataset**

```{r filter-seastar}
seastar <- rcca_data |>
  filter(str_detect(Classcode, "star")) |>
  arrange(Year)

seastar
```

```{r arrange-batstar}
batstar <- seastar |>
  filter(Classcode == "bat star") |>
  arrange(Year)

batstar
```


```{r}
# Original data saved 
write.csv(batstar_bwAvg, 
          "~/Documents/STATS_210_Spring_25/Project/project_part2/batstar.csv", 
          row.names = FALSE)
```



```{r select-site1}
batstar_bw <- batstar |>
  select(Site, Year, SurveyDate, Classcode, Amount) |>
  filter(Site == "Breakwater")

batstar_bw
```


```{r summarise-avg1}
batstar_bwAvg <- batstar_bw |>
  group_by(Site, Year) |>
  summarise(AverageAmount = mean(Amount, na.rm = TRUE), .groups = "drop")

# View the summarized data
batstar_bwAvg
```

```{r select-site2}
batstar_mb <- batstar |>
  select(Site, Year, SurveyDate, Classcode, Amount) |>
  filter(Site == "MacAbee")

batstar_mb
```

```{r summarise-avg2}
batstar_mbAvg <- batstar_mb |>
  group_by(Site, Year) |>
  summarise(AverageAmount = mean(Amount, na.rm = TRUE), .groups = "drop")

# View the summarized data
batstar_mbAvg
```

```{r}
batstar_all <- bind_rows(batstar_bwAvg, batstar_mbAvg) |>
  arrange(Year)

batstar_all
```

```{r clean-data}
# Clean data set number 1:
batstar_bwAvg

# Clean data set number 2:
batstar_mbAvg

# Clean data set number 1:
batstar_all
```


```{r save-data}
# Data set number 1:
write.csv(batstar_bwAvg, 
          "~/Documents/STATS_210_Spring_25/Project/project_part2/data-clean/batstar_bwAvg.csv", 
          row.names = FALSE)

# Data set number 2:
write.csv(batstar_bwAvg, 
          "~/Documents/STATS_210_Spring_25/Project/project_part2/data-clean/batstar_mbAvg.csv", 
          row.names = FALSE)

# Data set number 3:
write.csv(batstar_bwAvg, 
          "~/Documents/STATS_210_Spring_25/Project/project_part2/data-clean/batstar_all.csv", 
          row.names = FALSE)
```



## Part 2 - Data Visualizations


**Graphs for Bat Star count at Breakwater**

```{r graph-Breakwater}
# Graph One
batstar_bwAvg |> ggplot(aes(x = Year, y = AverageAmount)) + 
  geom_point() + 
  geom_smooth(fill = "steelblue") + 
  labs(x = "Years", 
       y = "Average Count", 
       title = "Average Amount of Bat Stars at Breakwater from 2006 to 2019") + 
  scale_x_continuous(breaks = seq(min(2006), max(2019), by = 1))

# Graph Two
batstar_bwAvg |> ggplot(aes(x = Year, y = AverageAmount)) + 
  geom_line(color = "steelblue") + 
  geom_point() + 
  labs(x = "Years", 
       y = "Average Count", 
       title = "Average Amount of Bat Stars at Breakwater from 2006 to 2019") + 
  scale_x_continuous(breaks = seq(min(2006), max(2019), by = 1))

# Graph Three
batstar_bwAvg |> ggplot(aes(x = Year, y = AverageAmount)) + 
  geom_col(fill = "steelblue") + 
  labs(x = "Years", 
       y = "Average Count", 
       title = "Average Amount of Bat Stars at Breakwater from 2006 to 2019") + 
  scale_x_continuous(breaks = seq(min(2006), max(2019), by = 1)) + 
  scale_y_continuous(breaks = seq(min(0), max(100), by = 10))
```


**Graphs for Bat Star count at MacAbee**

```{r graph-MacAbee}
# Graph One
batstar_mbAvg |> ggplot(aes(x = Year, y = AverageAmount)) + 
  geom_point() + 
  geom_smooth(fill = "aquamarine4") + 
  labs(x = "Years", 
       y = "Average Count", 
       title = "Average Amount of Bat Stars at MacAbee from 2006 to 2019") + 
  scale_x_continuous(breaks = seq(min(2006), max(2019), by = 1))

# Graph Two
batstar_mbAvg |> ggplot(aes(x = Year, y = AverageAmount)) + 
  geom_line(color = "aquamarine4") + 
  geom_point() + 
  labs(x = "Years", 
       y = "Average Count", 
       title = "Average Amount of Bat Stars at MacAbee from 2006 to 2019") + 
  scale_x_continuous(breaks = seq(min(2006), max(2019), by = 1))

# Graph Three
batstar_mbAvg |> ggplot(aes(x = Year, y = AverageAmount)) + 
  geom_col(fill = "aquamarine4") + 
  labs(x = "Years", 
       y = "Average Count", 
       title = "Average Amount of Bat Stars at MacAbee from 2006 to 2019") + 
  scale_x_continuous(breaks = seq(min(2006), max(2019), by = 1)) + 
  scale_y_continuous(breaks = seq(min(0), max(100), by = 10))
```


**Graphs for Bat Star count for both sites**

```{r graphs-all}
# Graph One
batstar_all |> ggplot(aes(x = Year, y = AverageAmount, color = Site)) + 
  geom_line() + 
  geom_point() + 
  labs(x = "Years", 
       y = "Average Count", 
       title = "Site Comparison for the Average Amount of Bat Stars from 2006 to 2019") + 
  scale_x_continuous(breaks = seq(min(2006), max(2019), by = 1))

# Graph Two
batstar_all |> ggplot() + 
  geom_smooth(mapping = aes(x = Year, y = AverageAmount, color = Site)) +
  labs(x = "Years", 
       y = "Average Count", 
       title = "Site Comparison for the Average Amount of Bat Stars from 2006 to 2019") + 
  scale_x_continuous(breaks = seq(min(2006), max(2019), by = 1))

# Graph Three
ggplot(batstar_all, aes(x = Site, y = AverageAmount)) + 
  geom_boxplot() + 
  labs(x = "Site", 
       y = "Average Count", 
       title = "Site Comparison for the Total Average Amount of Bat Stars")
```








